\textbf{ Spectrophotometric parallaxes with \project{The Cygnet}:\\ Accurate distances for luminous red-giant stars }

### 2018-08-07 paper/sp.tex

% To-do items
% -----------
% - draft discussion
% - make sure we address the sensibleness and generality of the linear model for the photometric space
% - Coryn's point that we should carefully explain how our model is geometric.
% - related: HWR's point that this is standardizeable candles, like the Ia SNe results.
% - related: Put in an explanation of WHY this works so well.
% - related: Put in an explanation of how it is that we aren't really modeling luminosity or distance but really just generating parallaxes.
% - does the model need a cute name, like The Swan?
% - check with WISE experts that we have the correct WISE photometry
% - we need an uncertainty estimate or noise model for our outputs.
% - we have set lambda = 30 by x-val.

@@

% text macros
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\equationname}{equation}
\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
\newcommand{\foreign}[1]{\textsl{#1}}
\newcommand{\etal}{\foreign{et al.}}
\newcommand{\eg}{\foreign{e.g.}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\wise}{\project{\acronym{WISE}}}
\newcommand{\zmass}{\project{\acronym{2MASS}}}
\newcommand{\sdssiv}{\project{\acronym{SDSS-IV}}}
\newcommand{\sdssv}{\project{\acronym{SDSS-V}}}

@@

\section{Assumptions of the method}

@@

There are not many choices
for building a model of the stars that is both justifiable probabilistically
and consistent with the assumptions stated in the previous Section.
Here we lay out the model and methodology.
We apply the method to real data in the following Sections.

@@

What did we do and what's cool?

Why did we make the assumptions we made and what did they gain us and cost us?

One strong assumption of this work is that our set of features is sufficient to
build a predictive model for distance.
Our arguments for our included features is that they ought to be sufficient to
estimate a dust-independent apparent magnitude and the stellar parameters (especially
age and surface gravity) that are sensitive to luminosity.
The derived model parameters shown in \figurename~\ref{fig:spectral} do indicate that
the model is capturing some or all of this information.
However, our choice of features is very rigid:
In principle the data themselves should tell us what features to include.
In that direction, it would make sense to choose features using a more complex
technique like deep learning or an auto-encoder (CITE).
These methods generate good features automatically.
However, they also involve an enormous set of hyper-parameter-like choices,
and they involve sacrificing certain kinds of interpretability.
However, we expect that a better set of features would do better; in this sense
the project here is just a first step towards precise spectrophotometric distance
estimates.

@@

Complain about our lack of a noise model.

(Mention independence of every result from the \gaia\ astrometry.)

What is this useful for and how do you use them safely?
HOGG: Show a map of the kinematic rotation of the Galaxy as a function of
x and y, colored by metallicity in \figurename~\ref{fig:disk}.
We see the kinematic center of the Galaxy! We see the rotation curve.
We see the chemical gradients.
\begin{figure}
\caption{A map of the kinematics and element abundances in the Milky Way disk.
  This map shows HOGG SOMETHING.\label{fig:disk}}
\end{figure}

How do you build a proper probabilistic model using these? How is the answer to that
related to the answer for \gaia\ itself?

>>>>

% To-do items
% -----------
% - draft discussion
% - make sure we address the sensibleness and generality of the linear model for the photometric space
% - Coryn's point that we should carefully explain how our model is geometric.
% - related: HWR's point that this is standardizeable candles, like the Ia SNe results.
% - related: Put in an explanation of WHY this works so well.
% - related: Put in an explanation of how it is that we aren't really modeling luminosity or distance but really just generating parallaxes.
% - does the model need a cute name, like The Swan?
% - check with WISE experts that we have the correct WISE photometry
% - we need an uncertainty estimate or noise model for our outputs.
% - we have set lambda = 50 by x-val. We could go as high as 60 without hurting our results.
% - We probably should cut in color-color space. Probably in J-K vs B-R and in H-W2.
% - We will match axes on relevant plots.
% - ...

@@

% text macros
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\sectionname}{Section}
\newcommand{\equationname}{equation}
\newcommand{\code}[1]{\texttt{\detokenize{#1}}}
\newcommand{\foreign}[1]{\textsl{#1}}
\newcommand{\etal}{\foreign{et al.}}
\newcommand{\eg}{\foreign{e.g.}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\wise}{\project{\acronym{WISE}}}
\newcommand{\zmass}{\project{\acronym{2MASS}}}
\newcommand{\sdssiv}{\project{\acronym{SDSS-IV}}}
\newcommand{\sdssv}{\project{\acronym{SDSS-V}}}

@@

\section{Assumptions of the method}\label{sec:assumptions}

@@

There are not many choices
for building a model of the stars that is both justifiable probabilistically
and consistent with the assumptions stated in the previous \sectionname.
Here we lay out the model and methodology.
We apply the method to real data in the following {\sectionname s}.

@@

This project demonstrates that photometry and spectroscopy can be used to predict
or estimate distances to stars.
This isn't new!
However, what's new is that it is possible to do this very well but with a purely
linear model acting on magnitudes and spectral (logarithmic) fluxes, and entirely
data-driven.
That is, a linear model trained on noisy parallax data from \gaia.

The model and method is based on a raft of assumptions, listed above
in \sectionname~\ref{sec:assumptions}.
We make some comments there, but we emphasize some particulars here:

One strong assumption of this work is that our set of features is sufficient to
build a predictive model for distance.
Our arguments for our included features is that they ought to be sufficient to
estimate a dust-independent apparent magnitude and the stellar parameters (especially
age and surface gravity) that are sensitive to luminosity.
In particular, since our model is linear in the logarithm of the parallax, it is
also linear in any kind of apparent magnitude at fixed distance and also linear
in distance modulus at fixed luminosity.
Our feature decisions were highly motivated by these kinds of considerations.
And the derived model parameters shown in \figurename~\ref{fig:spectral} do indicate that
the model is capturing some or all of the expected information in the spectra, at least.

However, our choice of features is very rigid:
In principle the data themselves should tell us what features to include.
In that direction, it would make sense to choose features using a more complex
technique like deep learning or an auto-encoder (CITE).
These methods generate good features automatically.
However, they also involve an enormous set of hyper-parameter-like choices,
and they involve sacrificing certain kinds of interpretability.
However, we expect that a better set of features would do better; in this sense
the project here is just a first step towards precise spectrophotometric distance
estimates.

@@

One significant limitation of the model is that it does not produce a sensible or
useful noise estimate or error bar on each spectrophotometric parallax estimate.
Although the linear model does permit simple noise propagation, that noise propagation
seriously under-estimates the scatter between the spectrophotometric parallaxes
and the best of the \gaia\ \acronym{DR2} astrometric parallaxes.
We thus used an empirical noise estimate to get our HOGG DETAIL HERE.
In principle it might have been better to build a noise model for our outputs,
and learned that noise model along with the coefficients.

Because of our train-and-test framework, each star is given a spectrophotometric
parallax based on coefficients derived from training on a complementary Training Set sample, of
which that star is not a member.
This framework ensures that the \gaia\ parallax $\gparallax_m$ of star $m$ is never used,
even indirectly, in generating the spectrophotometric parallax $\sparallax_m$ of star $m$.
That means that the parallax estimates generated here and given in \tablename~\ref{tab:output}
is statistically independent of the \gaia\ parallaxes and can be combined with them in
na\"ive ways, for example in inverse-variance-weighted averages.

What is this useful for and how do you use them safely?
HOGG: Show a map of the kinematic rotation of the Galaxy as a function of
x and y, colored by metallicity in \figurename~\ref{fig:disk}.
We see the kinematic center of the Galaxy! We see the rotation curve.
We see the chemical gradients. HOGG WHY CAN YOU NOT JUST READ OFF THINGS FROM THIS PLOT?
\begin{figure}
\caption{A map of the kinematics and element abundances in the Milky Way disk.
  This map shows the rotation of the Galaxy out to large radius, and also
  indicates the kinematic center of the disk rotation. Interpretation of
  the quantitative information in this plot requires care (see text).\label{fig:disk}}
\end{figure}

\figurename~\ref{fig:disk} shows that this sample of spectrophotometric parallaxes is useful for
mapping the Milky Way.
How does an investigator use these spectrophotometric parallaxes wisely?
How do you build a proper probabilistic model using these?
The answer is similar to the answer to the same question for the \gaia\ \acronym{DR2} data
themselves (\citealt{gaialf}): A likelihood function must be constructed for each of the
spectroscopic parallaxes, given assumptions about the noise in those values.
The sensible likelihood function for star $m$ would be HOGG PUT HERE.
Model fitting can then be built on those likelihood functions.
